Optimal model-based design of experiments for parameter precision: Supercritical Extraction case

### 2025-10-04 Article_Design_of_Experiment-Chamomile.tex
04.10

@@

\section{Introduction}

%\subfile{Sections/introduction_imp}
Supercritical CO2 extraction is a green extraction method that uses carbon dioxide in a supercritical state (above its critical point of 31.1 $^\circ$C and 73.8 bar) to extract compounds from materials. This is one of the most popular applications of supercritical fluids, which was described by many researchers, for example \citet{Sodeifian2017}, \citet{Reverchon1993} and \citet{Sovova1994}. Traditional methods, such as distillation and organic solvent extraction, are commonly employed although they have drawbacks. Distillation involves high temperatures that can lead to the thermal degradation of heat-sensitive compounds. This limitation has increased the popularity of alternative techniques, such as supercritical fluid extraction. Supercritical CO$_2$ is attractive due to its distinctive properties: it is inflammable, non-toxic and non-corrosive. Supercritical fluids can exhibit both gas- and liquid-like properties, allowing for adjustable dissolving power through changes in operating conditions.

@@

@@

		The residuals between the predicted and observed values were used to compute an unbiased estimate of the error variance. This variance, combined with the inverse of the design matrix's Gram matrix, yielded the covariance matrix of the coefficient estimates. The standard deviations (standard errors) of the coefficients were then obtained as the square roots of the diagonal elements of this covariance matrix. The variance-covariance is presented in Table \ref{tab:Uncertainty}.

@@

@@

		{\footnotesize 
			\begin{equation}
				\begin{aligned}
					\Xi^* &= \arg \min_{T^{\text{in}}(t), F(t)} 
					\int_{t_0}^{t_f} \left[ - \ln j_D(\Xi, \dot{x}) + 
					\frac{du(t)^\top}{dt} \, \bar{R} \, \frac{du(t)}{dt} \right] dt, \\
					\text{subject to:} \quad 
					& \dot{x} = G(x, t, \theta; \Xi), \\
					& t_0 = 0 \, \text{min}, \quad t_f = 600 \, \text{min}, \\
					& T^0 = T^{\text{in}}(t=0), \\
					& P(t) \in \{100, 125, 150, 175, 200\} \, \text{bar}, \\
					& 31^\circ\text{C} \leq T^{\text{in}}(t) \leq 40^\circ\text{C}, \\
					& 3.33 \cdot 10^{-5} \leq F(t) \leq 6.67 \cdot 10^{-5}, \\
					& \mathcal{R} = 
					\begin{bmatrix} 
						0.1 & 0 \\ 
						0 & 0.01 
					\end{bmatrix}, \\
					& \frac{du(t)}{dt} = 
					\begin{bmatrix} 
						\dfrac{dF(t)}{dt} \\ 
						\dfrac{dT^{\text{in}}(t)}{dt} 
					\end{bmatrix}.
				\end{aligned}
				\label{EQ:Formulation_2}
		\end{equation} }

@@

		The results and observations discussed in this article are local and should be considered valid only for the model presented in \citet{Sliczniuk2024}. Although general conclusions about the curvature of the yield curve are independent of the process model, a completely different set of control profiles should be expected if the same technique is applied to a different model. Such strong dependence on the model is one of the most significant limitations of this work. Moreover, the model predicts the data set in a region which has not been validated during previous experiments. On the one hand, this leaves room for model validation, while on the other hand, the obtained set of data points might differ from the predictions.

@@

\clearpage
%\newpage

@@

\clearpage
\onecolumn
\appendix
\section{Literature Review} \label{CH: Literature}
Table \ref{fig: SFE_Literature} provides a list of studies on SFE including the modeling part presented from 2014 onwards

\begin{table*}[h!]
	\centering
	\resizebox{\textwidth}{!}{%
	\begin{tabular}{|p{2.5cm}|p{2cm}|p{1.75cm}|p{4.5cm}|p{1.5cm}|p{1.5cm}|p{3cm}|}
		\hline
		\textbf{Reference} & \textbf{Substrate} & \textbf{Solvent} & \textbf{Model} & \textbf{T (°C)} & \textbf{P (MPa)} & \textbf{Flow-rate of CO$_2$} \\
		\hline
		\citet{Botelho2015} & Black sesame seed & CO$_2$ & BIC (\citet{Sovova1994}) \newline Local adsorption equilibrium model (\citet{Goto1993}) &  40, 60 & 20-40 & 5.9$\times 10^{-5}$ [kg/s]\\ \hline
		\citet{Melo2014} & Eucalyptus globulus bark & CO$_2$, \newline Co-solvent: ethanol & Diffusion model (\citet{Crank1975}) \newline \citet{Brunner1994} \newline Desorption model (\citet{Cocero2001}) \newline SSP model (\citet{Gaspar2003})&  40 & 20 & 2.78$\times 10^{-3}$ [kg/s]\\ \hline
		\citet{Liu2014} & Calycopteris floribunda leaves & CO$_2$, \newline Co-solvent: ethanol & BIC (\citet{Sovova1994}) \newline Diffusion model (\citet{Crank1975}) &  30–50 & 10-30 & 4.17$\times 10^{-4}$ [kg/s]\\ \hline
		\citet{Tomita2014} & Rice bran & CO$_2$ & Partitioning coefficient model (\citet{Kubatova2002}) & 40–80 & 20-40 & 0.17-1.5$\times 10^{-4}$ [L/s]\\ 
		\hline
	\end{tabular} }
	\caption{SFE studies encompassing modeling part}
	\label{fig: SFE_Literature}
\end{table*}

>>>>

\usepackage[switch]{lineno}

@@

\section{Introduction}
\linenumbers
%\subfile{Sections/introduction_imp}
Supercritical CO2 extraction is a green extraction method that uses carbon dioxide in a supercritical state (above its critical point of 31.1 $^\circ$C and 73.8 bar) to extract compounds from materials. This is one of the most popular applications of supercritical fluids, which was described by many researchers, for example \citet{Sodeifian2017}, \citet{Reverchon1993} and \citet{Sovova1994}. Traditional methods, such as distillation and organic solvent extraction, are commonly employed although they have drawbacks. Distillation involves high temperatures that can lead to the thermal degradation of heat-sensitive compounds. This limitation has increased the popularity of alternative techniques, such as supercritical fluid extraction. Supercritical CO$_2$ is attractive due to its distinctive properties: it is inflammable, non-toxic and non-corrosive. Supercritical fluids can exhibit both gas- and liquid-like properties, allowing for adjustable dissolving power through changes in operating conditions.

@@

The aforementioned methods, along with many others, have been commonly employed for modeling the SFE process. Appendix \ref{CH: Literature} provides a list of studies on SFE, including modeling approaches, raw materials, and operating conditions.

@@

		The residuals between predicted and observed values were used to compute an unbiased estimate of the error variance. This variance, combined with the inverse of the design matrix's Gram matrix, yielded the covariance matrix of the coefficient estimates. The standard deviations (standard errors) of the coefficients were then obtained as the square roots of the diagonal elements of this covariance matrix. The variance-covariance is presented in Table \ref{tab:Uncertainty}.

@@

		The measurement error can be obtained based on the accuracy of the measuring device, or empirically from the spread of data around the model output. The first approach can be utilised only in the absence of external error sources. Given the imperfections of the experimental setup, such as delayed measurements and residuals of the product in pipes and the draining vessel, it was decided that the second approach is more suitable for this work. Based on the \citet{Sliczniuk2024}, the empirical measurement error was estimated to be 0.03.

@@

		{\footnotesize 
			\begin{equation}
				\begin{aligned}
					\Xi^* &= \arg \min_{T^{\text{in}}(t), F(t)} 
					\int_{t_0}^{t_f} \left[ - \ln j_D(\Xi, \dot{x}) + 
					\frac{du(t)^\top}{dt} \, \bar{R} \, \frac{du(t)}{dt} \right] dt, \\
					\text{subject to:} \quad 
					& \dot{x} = G(x, t, \theta; \Xi), \\
					& t_0 = 0 \, \text{min}, \quad t_f = 600 \, \text{min}, \\
					& T^0 = T^{\text{in}}(t=0), \\
					& P(t) \in \{100, 125, 150, 175, 200\} \, \text{bar}, \\
					& 30^\circ\text{C} \leq T^{\text{in}}(t) \leq 40^\circ\text{C}, \\
					& 3.33 \cdot 10^{-5} \leq F(t) \leq 6.67 \cdot 10^{-5}, \\
					& \mathcal{R} = 
					\begin{bmatrix} 
						\mathcal{R}_{FF} & \mathcal{R}_{FT} \\ 
						\mathcal{R}_{TT} & \mathcal{R}_{TF} 
					\end{bmatrix}=
					\begin{bmatrix} 
						0.1 & 0 \\ 
						0 & 0.01 
					\end{bmatrix}, \\
					& \frac{du(t)}{dt} = 
					\begin{bmatrix} 
						\dfrac{dF(t)}{dt} \\ 
						\dfrac{dT^{\text{in}}(t)}{dt} 
					\end{bmatrix}.
				\end{aligned}
				\label{EQ:Formulation_2}
		\end{equation} }

@@

		To the authors knowledge, model-based dynamic design of experiments has not previously been applied to supercritical extraction, hence the authors cannot directly compare the obtained results against the literature. Nevertheless, some researchers proposed running an extraction process under dynamically changing operating conditions. For instance, \citet{Simandi1998} proposed a fractionated extraction of oregano with a stepwise increase in pressure (80, 120, 200, and 300 bar). As a result, the obtained yield curve exhibits non-monotonic patterns. \citet{Hiraga2025} suggested a pressure swing CO$_2$ extraction for the decaffeination of green coffee beans. The authors varied the pressure up and down during the operation, achieving a high degree of decaffeination with a reduced CO$_2$ supply. As a result, nearly 100\% decaffeination was achieved at 353 K and 30 MPa, using 80\% of the CO$_2$ supply compared to the case of constant pressure operation. Some of the extraction yield plots presented by \citet{Hiraga2025} are characterised by non-monotonic behaviour. It can be concluded that non-monotonic extraction curves have been reported in the literature.

		The results and observations discussed in this article are local and should be considered valid only for the model presented in \citet{Sliczniuk2024}. Although general conclusions about the curvature of the yield curve are independent of the process model, a completely different set of control profiles should be expected if the same technique is applied to a different model. Such strong dependence on the model is one of the most significant limitations of this work. Moreover, the model predicts the data set in a region which has not been validated during previous experiments. This leaves room for model validation, while on the other hand, the obtained set of data points might differ from the predictions.

@@

%\clearpage
%\newpage

@@

\clearpage
\onecolumn
\appendix
\section{Literature Review} \label{CH: Literature}
Table \ref{fig: SFE_Literature} provides a list of studies on SFE including the modelling part presented from 2014 onwards

\begin{table*}[h!]
	\centering
	\resizebox{\textwidth}{!}{%
	\begin{tabular}{|p{2.5cm}|p{2cm}|p{1.75cm}|p{4.5cm}|p{1.5cm}|p{1.5cm}|p{3cm}|}
		\hline
		\textbf{Reference} & \textbf{Substrate} & \textbf{Solvent} & \textbf{Model} & \textbf{T (°C)} & \textbf{P (MPa)} & \textbf{Flow-rate of CO$_2$} \\
		\hline
		\citet{Botelho2015} & Black sesame seed & CO$_2$ & BIC (\citet{Sovova1994}) \newline Local adsorption equilibrium model (\citet{Goto1993}) &  40, 60 & 20-40 & 5.9$\times 10^{-5}$ [kg/s]\\ \hline
		\citet{Melo2014} & Eucalyptus globulus bark & CO$_2$, \newline Co-solvent: ethanol & Diffusion model (\citet{Crank1975}) \newline \citet{Brunner1994} \newline Desorption model (\citet{Cocero2001}) \newline SSP model (\citet{Gaspar2003})&  40 & 20 & 2.78$\times 10^{-3}$ [kg/s]\\ \hline
		\citet{Liu2014} & Calycopteris floribunda leaves & CO$_2$, \newline Co-solvent: ethanol & BIC (\citet{Sovova1994}) \newline Diffusion model (\citet{Crank1975}) &  30–50 & 10-30 & 4.17$\times 10^{-4}$ [kg/s]\\ \hline
		\citet{Tomita2014} & Rice bran & CO$_2$ & Partitioning coefficient model (\citet{Kubatova2002}) & 40–80 & 20-40 & 0.17-1.5$\times 10^{-4}$ [L/s]\\ \hline
		\citet{Kupski2017} & Hop pellet & CO$_2$ & BIC (\citet{Sovova1994}) \newline \citet{Reverchon1996} &  35-55 & 10-20 & 3.25$\times 10^{-5}$ [kg/s]\\ \hline
		\citet{Pavlic2017} & Sage & CO$_2$ & \citet{Reverchon1996} \newline  \citet{Brunner1994} \newline \citet{Kandiah1990} &  40-60 & 10-30 & 5.56-11.1$\times 10^{-5}$ [kg/s]\\ \hline
		\citet{Sartori2017} & Banana & CO$_2$ & \citet{Sovova1994} \newline  \citet{Reverchon1994} \newline \citet{Esquivel1999} &  40-80 & 20-50 & 1.01$\times 10^{-4}$ [L/s]\\ \hline
		\citet{Hall2018} & Papaya seeds & CO$_2$ & \citet{Sovova1994} \newline  \citet{Goto1993} \newline \citet{Esquivel1999} &  40 & 15-30 & 6.11-19.4$\times 10^{-5}$ [kg/s]\\ \hline
		\citet{Putra2018} & Peanut skin & CO$_2$, \newline Co-solvent: ethanol & \citet{Brunner1994} \newline \citet{Esquivel1999} &  40-70  & 10-30 & 5$\times 10^{-5}$ [L/s]\\ \hline
		\citet{Putra2018} & Wheat germ & CO$_2$ & \citet{Brunner1994} \newline \citet{Esquivel1999} \newline \citet{Kandiah1990} \newline \citet{Reverchon1994} &  40-60  & 25-35 & 0.56 - 1.11 $\times 10^{-4}$ [kg/s]\\ \hline
		\citet{Pavlic2020} & Raspberry seed & CO$_2$ & \citet{Brunner1994} \newline \citet{Esquivel1999} \newline \citet{Kandiah1990} \newline \citet{Reverchon1994} &  40-60  & 25-35 & 0.56 - 1.11 $\times 10^{-4}$ [kg/s]\\ \hline
		\citet{Priyanka2020} & Carrot seed & CO$_2$, \newline Co-solvent: ethanol & BIC (\citet{Sovova1994}) \newline \citet{Reverchon1997} &  50-70  & 20-40 & 0.83 - 2.5 $\times 10^{-4}$ [kg/s]\\ \hline
		\citet{Casas2021} & Acacia dealbata - ﬂowers & CO$_2$, \newline Co-solvent: ethanol & \citet{Esquivel1999} \newline \citet{Sovova1994} &  35-55  & 10-35 & 4.17 $\times 10^{-4}$ [kg/s]\\ \hline
		\citet{Acosta2014} & Wood (\textit{Mezilaurus itauba}) & CO$_2$, \newline Co-solvent: ethanol (2\%, 5\%, 8\%) & BIC(\citet{Sovova1994}) &  38-52  & 8-22 & 5 $\times 10^{-5}$ [L/s]\\ \hline
		\citet{Zekovic2014} & \textit{Ocimum basilicum L.} & CO$_2$ & \citet{Brunner1994} \newline \citet{Kandiah1990} \newline \citet{Esquivel1999} &  40-60  & 10-30 & 2.71 $\times 10^{-2}$ [L/s]\\ \hline
	\end{tabular} }
	\caption{SFE studies encompassing modeling part}
	\label{fig: SFE_Literature}
\end{table*}
